---
title: "Tidy Tuesday: 2021 Week 14 - Makeup Shades"
author: "Nelson Pray"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r initial, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,  fig.width = 12, fig.height = 9.2)
```

## Introduction

I found this week's dataset a little harder to work with than last week's. I had a lot of ideas of what I wanted to create, inspired by [many](https://twitter.com/SarinaSinghK/status/1377329518009065479) [awesome](https://twitter.com/quite_grey/status/1377374520168882180) [examples](https://twitter.com/li_bennett_/status/1377289004022960129) but I struggled to get my plots to work correctly. I almost gave up but decided to keep it simple and went with a more basic plot than I initially planned.

One of my breakthrough moments came when I decided to check which brands were the top selling brands in the world. I found a table of the top 50 cosmetic brands of 2020 from [Brand Finance](https://brandirectory.com/rankings/cosmetics/table) and went through the brands in the dataset and found the top matching ones.

## Setup

The setup for this plot was pretty straight forward. Although this was my first time using the {showtext} package and I must say it worked perfectly and I was very happy with it

```{r setup, message=FALSE, warning=FALSE}
# import libraries
library(tidyverse)
library(showtext)
library(ggtext)

# add custom fonts
font_add_google("Merienda", "main_font")
font_add_google("Merriweather Sans", "secondary_font")

showtext_opts(dpi = 300)
showtext_auto()

# define colours
grey <- "#5A5A66"
dark_text <- "#3A2115"   # found the hex matching min(lightness) of all shades
light_text <- "#FEFEFE"  # found the hex matching max(lightness) of all shades
avg_colour <- "#CF9B74"  # found the closest hex matching mean(lightness) of all shades
```

## Tidying

When I first starting working with this dataset I ran into an interesting error while using the `tt_load()` function from the {tidytuesdayR} package. I think I must have been requesting the data too many times as I requested the data each time I ran the cell that there was a message about github and needing to wait. That is why I ended up using the `read_csv()` function from {readr}.

```{r tidying, message=FALSE, warning=FALSE}
# read in data
all_shades<- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allShades.csv')

# define the top companies
top_companies <- sort(c("L'Oréal", "Estée Lauder", "Clinique", "Guerlain", "Shiseido",
                   "MAC", "Maybelline", "Lancôme", "Benefit Cosmetics"), decreasing = TRUE)

# grab the top companies, the hex code for the colour, and the lightness
df <- all_shades %>% 
  filter(brand %in% top_companies) %>% 
  select(brand, hex, lightness)

df$brand <- as.factor(df$brand)

df <- df %>% 
  group_by(brand) %>% 
  mutate(mean_lightness = mean(lightness)) %>% 
  ungroup() %>% 
  mutate(brand = fct_reorder(brand, mean_lightness))
```

## Plotting

As mentioned above, I had quite a few issues when I originally started working with the data for this week. I had plans of making [donut plots](https://www.r-graph-gallery.com/doughnut-plot.html) but I couldn't get my data in the right format to work properly. That is why I went with the simpler scatter plot. Another issue I had while making the finished plot is getting the colours correct, I still don't think they are perfect because some of the shades are hard to see on the grey background, but alas I could probably spend days just tweaking the colour and still not end up with something I am completely happy with.

```{r plot, message=FALSE, warning=FALSE}
ggplot(df, aes(brand, lightness, colour = hex)) +
  geom_jitter(width = 0.15, 
              height = 0.0, 
              size = 2) +
  scale_colour_identity() +
  labs(title = "Shades of Bias",
       subtitle = "For decades, the beauty industry has had issues with racial bias when it comes to <br/>their products. Below are the palettes of some of the best selling makeup brands in <br/>the world. We can clearly see that there are many more lighter shades than there <br/>are darker shades of makeup.", 
       x = "Brands in order of average lightness<br/>
       <span style = 'color:#3A2115;'><-Dark</span>
       <span style = 'color:#CF9B74;'>to</span> 
       <span style = 'color:#FEFEFE;'>Light -></span>",
       y = "<span style = 'color:#3A2115;'><-Dark</span> to <span style = 'color:#FEFEFE;'>Light -></span>",
       caption = "Data: The Pudding, Brand Finance | Graphic: @nelsonpray") +
  scale_x_discrete(guide = guide_axis(angle = -45)) +
  theme_minimal() +
  theme(panel.background = element_rect(colour = grey, 
                                        fill = grey),
        panel.grid = element_blank(),
        plot.background = element_rect(colour = grey, 
                                       fill = grey),
        plot.margin = margin(.1, .6, 1.1, .6, "cm"),
        plot.title = element_markdown(family = "main_font",
                                      face = "bold",
                                      size = 24,
                                      colour = light_text,
                                      padding = unit(c(20, 3, 0, 0), "pt")),
        plot.subtitle = element_markdown(family = "secondary_font",
                                      size = 10,
                                      colour = light_text),
        plot.caption = element_markdown(family = "secondary_font", 
                                        colour = light_text, 
                                        size = 8, 
                                        padding = unit(c(10, 3, 3, 3), "pt")),
        axis.text.y = element_blank(),
        axis.text = element_markdown(family = "secondary_font",
                                     size = 10,
                                     colour = light_text),
        axis.title.x = element_markdown(family = "main_font",
                                        size = "16",
                                        colour = light_text),
        axis.title.y = element_markdown(family = "main_font",
                                        size = "16",
                                        colour = avg_colour),
        axis.line.x.top = element_blank(),
        axis.line.x.bottom = element_line(colour = "#8c99a5"))
```

## Conclusion

This week was a bit of a struggle for me when it came to creating the plot. Having seen how awesome other people's plots were I had a serious case of imposter syndrome and almost stopped. Thankfully I perserved and finished the plot. It might not be my best work but it is still better than having no work at all.